Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-15 appendix.qmd
Updated appendix
Updated point alpha

trimesh_removed_pbmc <- ggplot() + 
  geom_point(
    data = umap_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.2,
    size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

algo_obj_pbmc <- fit_highd_model(
  highd_data = training_data_pbmc, 
  nldr_data = tsne_pbmc, 
  bin1 = num_bins_x_pbmc, 
  q = 0.1, 
  benchmark_highdens = 1)

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = tsne_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.2,
    size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by MSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters. The PBMC3k data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close. Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations. Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%

>>>>

trimesh_removed_pbmc <- ggplot() + 
  geom_point(
    data = umap_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.3#,
    #size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

algo_obj_pbmc <- fit_highd_model(
  highd_data = training_data_pbmc, 
  nldr_data = tsne_pbmc, 
  bin1 = num_bins_x_pbmc, 
  q = 0.1, 
  benchmark_highdens = 5)

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = tsne_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.3#,
    #size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by RMSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters. The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close. Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations. Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%

### 2025-05-15 appendix.tex
Updated appendix
Updated point alpha

\caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
(Figure~\ref{fig-pbmc-mse-umap} b) by MSE plot
(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
hyper-parameters. The PBMC3k data (\(n =  5858\)) has three clusters in
\(9\text{-}D\), where three clusters are close. Two \gD{} projection
from a tour on \(9\text{-}D\) of the model fit with
Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.06\),
\(b = 616/86 (22, 28)\)) shows three-well separated clusters with big
separations. On the other hand, the model fit with
Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.06\),
\(b = 704/227 (22, 32)\)) shows three-well separated clusters with small
separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
langevitour animations are available at
\url{https://youtu.be/0cKX_HG_n0k} and
\url{https://youtu.be/KhJvsRtaX04} respectively.}

>>>>

\caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
(Figure~\ref{fig-pbmc-mse-umap} b) by RMSE plot
(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
hyper-parameters. The PBMC data (\(n =  5858\)) has three clusters in
\(9\text{-}D\), where three clusters are close. Two \gD{} projection
from a tour on \(9\text{-}D\) of the model fit with
Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.06\),
\(b = 616/86 (22, 28)\)) shows three-well separated clusters with big
separations. On the other hand, the model fit with
Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.06\),
\(b = 704/227 (22, 32)\)) shows three-well separated clusters with small
separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
langevitour animations are available at
\url{https://youtu.be/0cKX_HG_n0k} and
\url{https://youtu.be/KhJvsRtaX04} respectively.}

### 2025-05-15 paper.qmd
Updated point alpha

trimesh_removed_pbmc <- ggplot() + 
  geom_point(
    data = umap_pbmc_scaled_with_cluster,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = cluster
    ),
    alpha = 0.2,
    size = 0.1#,
    #color = '#a65628'
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("a1", c(0.08, 0.9), cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

>>>>

trimesh_removed_pbmc <- ggplot() + 
  geom_point(
    data = umap_pbmc_scaled_with_cluster,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = cluster
    ),
    alpha = 0.2#,
    #size = 0.1#,
    #color = '#a65628'
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("a1", c(0.08, 0.9), cex = 2) +
  theme(
    aspect.ratio = 1
  )
```
